Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-04 methods_paper/fabl-paper.tex
update.

\section{Case Studies}
\label{sex:case-studies}

We now demonstrate the power of our method through a case study of documented identifiable deaths (DID) from the El Salvadoran Civil War. Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors. Secondly, the case study reveals the impact of independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.

@@

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualties, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casualty lists. To obtain a more accurate estimate of the
casualties then, we follow Sadinle's 2017 paper and link records of El
Salvadoran civilian casualties from two sources: El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

@@

We only utilized records with nonmissing entries for given and last
name, results in \(n_A = 4420\) files in CHDES and \(n_B = 1323\) files
in ERTL. The names were standardized to account for common misspellings
in the Spanish language, and then compared using a modified Levenstein
distance to account for the fact that second names are often omitted.
Place of birth is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, we excluded it from our linkage
process. Thus we conduct linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters.

	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP}

	\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
\end{figure}

To mirror the original implementation, we constructed the comparison
vectors using 4 levels of agreement for each field, according to the
thresholds provided in Figure XX. This took 422 seconds for our proposed
method and 240 for \texttt{BRL}. However, we observed that posterior
distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
parameters were nearly identical, and that of the
\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
realized in the data. This leads us to believe that such high number of
agreement levels creates unnecessary distinctions in the data and makes
the comparison vectors less interpretable. Therefore we re-ran our
analysis with fewer agreement levels for each field (see Figure XX), and
obtained analogous results. With 216 possible agreement patterns, 159
were realized in the data, and our proposed method became much faster,
finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
relatively unchanged from the first implementation. This demonstrates
the way that the computational complexity of our method depends on the
number of unique agreement patterns, and how significant computational
gains can be made by simplifying the construction of the comparison
vectors. We also note that estimates of the \(\mathbf{m}\) and
\(\mathbf{u}\) parameters under each method are very similar, as shown in
Figure~\ref{fig:m-and-u}.

>>>>

\section{Application to Civilian Casualties from the El Salvadoran Civil War}
\label{sex:case-studies}

In this section, we revisit an application to the El Salvadoran Civil War, where we estimate the number of documented identifiable deaths (DID). Though the data files used here are small, this study shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors. This case study also considers the ramifications of 
independently sampling $Z_j$ rather than strictly enforcing one-to-one matching as done by \texttt{BRL}.

@@

The country of El Salvador was immersed in civil war from 1980 to 1991,
and throughout the time, several organizations attempted to document
casualties of the conflict. When estimating the total number of
casualties, one cannot simply sum the numbers recorded by each
organization, as it is likely that the same individuals are recorded in
multiple casualty lists. Thanks to the Human Rights Data Analysis Group (HRDAG), we utilize 
two different sources, namely,  El Rescate - Tutela
Regal (ERTL) and the Salvadoran Human Rights Commission (CDHES, by its
acronym in Spanish). The ERTL dataset consists of digitized reports that
had been published throughout the conflict. The CDHES dataset consists
of casualties that had been reported directly to the organization, and
later digitized.

\textcolor{red}{Beka: These data sets need more information regarding how the data sets were collected as this is quite important. See Sadinle (2017) and references for info). }

@@

In this paper, we utilize records that have non-missing entries for given and last name, which results 
in \(n_1 = 4420\) files in CHDES and \(n_2 = 1323\) files
in ERTL. We standardized names to account for common misspellings
in the Spanish language. The string distance used is a modified Levenstein
distance to account for the fact that second names are often omitted.
Place of birth is recorded by municipality and department within that
municipality; however, since department was missing in 95\% of records
in CHDES and 80\% of records in ERTL, this was excluded from our analyses. 
Thus, we conduct record linkage using given name, last name,
municipality, and day, month, and year of death. We again use flat
priors for the \(\mathbf{m}\) and \(\mathbf{u}\) parameters. To our knowledge, we have followed the same guidelines/settings in \cite{sadinle_bayesian_2017}.

To mirror the original implementation of \cite{sadinle_bayesian_2017}, we construct the comparison
vectors using four levels of agreement for each field, according to the
thresholds provided in \textcolor{red}{Figure XX}. This took 422 seconds for our proposed
method, \texttt{fabl}, and 240 for \texttt{BRL}. However, we observed that posterior
distributions of several levels of the \(\mathbf{m}\) and \(\mathbf{u}\)
parameters were nearly identical, and that of the
\(4^5 \times 2 = 2048\) possible agreement patterns, only 1173 are
realized in the data. This leads us to believe that such high number of
agreement levels creates unnecessary distinctions in the data and makes
the comparison vectors less interpretable.

Therefore we re-ran our analysis with fewer agreement levels for each field \textcolor{red}{(see Figure XX)}, and
obtained analogous results. With 216 possible agreement patterns, 159
were realized in the data, and our proposed method, \texttt{fabl} became was faster,
finishing in 124 seconds. Meanwhile \texttt{BRL} took 239 seconds,
relatively unchanged from the first implementation. This demonstrates
the way that the computational complexity of our method depends on the
number of unique agreement patterns, and how significant computational
gains can be made by simplifying the construction of the comparison
vectors. The estimates of the \(\mathbf{m}\) parameters under each method are very similar, as shown in
Figure~\ref{fig:m-and-u}.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/m_posterior_smallP} 
\caption{Posterior estimates of m parameters with 95\% credible intervals}\label{fig:m-and-u}
\label{fig:m-and-u}
\end{center}
\end{figure}
